Biomass predictions and VOCC

### 2019-08-06 analysis/VOCC/4-vocc-by-spp.Rmd
linking vocc and species density code, still in development

```{r}
# # Species run so far...
 species <- "Arrowtooth Flounder"
 species <- "Pacific Cod"
 species <- "Sablefish"
# species <- "Silvergray Rockfish"
#  species <- "Lingcod"
#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
#  species <- "Quillback Rockfish"
#  species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
```

Region 
```{r}
# region <- "All synoptic surveys"
 region <- "Both odd year surveys"
 region <- "West Coast Vancouver Island"
# region <- "West Coast Haida Gwaii"
```

Make climate vectors for biannual changes between surveys
```{r}
trend <- FALSE
start_time <- 2015
end_time <- 2018

climate <- "temperature"
temp_threshold <- c(1)
optimal_temp <- 7 # yellow-eye 4

climate <- "do"
do_threshold <- c(0.5)
optimal_do <- 2.5 # yellow-eye 4

# climate <- "DO and temperature"
# max_thresholds <- c(Inf, 1)
# min_thresholds <- c(0.5, Inf)

@@

Run all subsequent code...
```{r global_options, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
#library(gfranges)

@@

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}
```

@@

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")
climate <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
```

Calculate climate vectors
```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate, "-",  
    min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
})

@@

saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate, "-",  
  min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
}

@@

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
```

Extract biomass values to match spatial and temporal range of climate vectors
```{r}
vocc_by_sp <- list()
for (i in seq_along(biomass_list)){

@@

if (length(variable_names) == 1) {
  vectors <- vocc %>%
    mutate(
      start_cell = icell, vector_id = paste0(id, "_to_", tid),
      initial_value = var_1_s, source_end_value = var_1_e, target_value = as.numeric(target_values)
    ) %>%
    filter(distance > 1)  %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan) 
  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))

} else {
  vectors <- vocc %>%
    mutate(
      start_cell = icell, vector_id = paste0(id, "_to_", tid),
      initial_value = s, source_end_value = e, target_value = target_values
    ) %>%
    filter(distance > 1) %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan) 
  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))

@@

@@

vocc_by_sp[[i]] <- rbind(vector_source, vector_target) %>% 
  filter(!is.na(icell)) %>% filter(source_density > 0.000001) %>% 
  mutate(perc_change = (end_density - start_density)/start_density, 
    cell_type = as.factor(cell_type)) %>% 
  rename(X = x, Y = y)

icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
mean.change <- mean(icells$perc_change, na.rm = TRUE)
sd.change <- sd(icells$perc_change, na.rm = TRUE)

@@

mean.climate <- mean(icells$end_climate, na.rm = TRUE)
sd.climate <- sd(icells$end_climate, na.rm = TRUE)

# glimpse(icells)
# range(icells$perc_change)
# range(vocc_by_sp[[i]]$std_change)

Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
Xmin <- quantile(vocc_by_sp[[i]]$X, 0.0005)
std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% mutate(std_change = (perc_change)/sd.change, 
  std_source_density = (source_density - mean.density)/(sd.density*2), 
  std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
  mean_change = mean.change, sd_change = sd.change,
  mean_source_density = mean.density, sd_density = sd.density,
  mean_climate = mean.climate, sd_climate = sd.climate,
  std_end_climate2 = std_end_climate^2, 
  distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, # set intercept at scale of cells
  distance2 = distance*distance
  ) %>% filter ( X >= Xmin & X <= Xmax )

# range(vocc_by_sp[[i]]$X)

#glimpse(vocc_by_sp[[i]])
}
names(vocc_by_sp) <- names(biomass_list)
```

@@

mature_plot2 <- ggplot(data, aes(distance, y = std_change, 
  group = cell_type, size = start_density, colour = end_climate)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y~x+I(x^2) )+
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_size_continuous(guide = "none") +
  # scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))

@@

mature_plot4 <- ggplot(filter(data, distance < 10), aes(source_density, y = std_change , 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

@@

@@

Boosted regression
```{r}
data <- vocc_by_sp[["adult"]]

@@

```{r}
data <- vocc_by_sp[["adult"]]

  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 100)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
  }

  sdmTMB::plot_spde(spde)
```

@@

@@

@@

itest <- lmerTest::lmer(std_change ~ 
      cell_type * std_source_density + 
      cell_type * distance + cell_type * distance2 +
      std_end_climate + std_end_climate2 + (1 | vector_id), 
      REML = F, data = idata)
summary(itest)
plot(itest)
```
```{r eval=FALSE}
library(lme4)
itest3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
    cell_type * distance + cell_type * distance2 + 
    std_end_climate + std_end_climate2 + (1 | vector_id), data = idata)
summary(itest3)
plot(itest3, ask = FALSE)
```

```{r}
mi <- gbm::gbm(
  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
  data = idata, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )

@@

```{r}
  if (region == "Both odd year surveys") {
    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 200)
  }

  if (region == "West Coast Vancouver Island") {
    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 100)
  }

  if (region == "West Coast Haida Gwaii") {
    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 60)
  }

  sdmTMB::plot_spde(ispde)
```

@@

varnames <- colnames(m$tmb_data$X_ij)
#b_j <- m$model$par
#source_intercept <- .sd[1,]
#target_intercept <- b_j[2]
#source_slope <- b_j[3]

.sd <- summary(m$sd_report)
.sd <- as.data.frame(.sd)
names(.sd) <- c("beta","se")
.sd$ci <- .sd$se*1.96
var <- varnames
var[1] <- "intercept"
.sd <- .sd[1:length(var), ]
row.names(.sd) <- NULL
.sd <-cbind(var, .sd)

.sd$species <- spp 
.sd$climate <- climate 
.sd$min_thresholds <- min_thresholds
.sd$max_thresholds <- max_thresholds
.sd$optimum <- optimal_value
.sd$start_time <- start_time
.sd$timespan <- delta_t_total

# intercept = mean std percent change in biomass for cell with avarage source biomass
data$mean_source_density[1]
range(data$source_density)
```

>>>>

```{r}
# # Species run so far...
 species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
#  species <- "Lingcod"
#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
#  species <- "Quillback Rockfish"
#  species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
```

Region 
```{r}
# region <- "All synoptic surveys"
 region <- "West Coast Vancouver Island"
# region <- "Both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

Make climate vectors for biannual changes between surveys
```{r}
trend <- FALSE
start_time <- 2008
end_time <- 2011

#climate <- "temperature"
temp_threshold <- c(0.75)
optimal_temp <- 6.5 # yellow-eye 4

#climate <- "do"
do_threshold <- c(0.25)
optimal_do <- 2 # yellow-eye 4

climate <- "do and temperature"
max_thresholds <- c(Inf, 0.75)
min_thresholds <- c(0.25, Inf)
optimal_value <- c(6.5, 2)
```

@@

Run all subsequent code...
```{r global_options, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

@@

# if (region == "West Coast Haida Gwaii") {
#   survey <- c("SYN WCHG")
#   model_ssid <- c(16)
#   ssid_string <- paste0(model_ssid, collapse = "n")
#   trend_indices_do <- c(1, 1, 1, 2, 2)
#   trend_indices_temp <- c(1, 1, 1, 2, 2)
#   years <- NULL
# }
```

@@

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")

optimal_string <- paste0(optimal_value, collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate_string)))
```

# Calculate climate vectors

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
})

@@

saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
}
# glimpse(vocc)
```

### Plot climate vectors
```{r}

if(climate== "temperature") {
vocc$diff <- vocc$units_per_decade/10*delta_t_total

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

if (climate == "do") {
vocc$diff <- vocc$units_per_decade/10*delta_t_total

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

if (climate == "do and temperature") {

vocc$temp_diff <- vocc$temp.units_per_decade/10*delta_t_total

vocc_plot_temp <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "temp_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)+ ggtitle("temperature")

vocc$do_diff <- vocc$do_est.units_per_decade/10*delta_t_total

vocc_plot_do <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "do_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
) + ggtitle("do")

plot <- gridExtra::grid.arrange(
  vocc_plot_temp, vocc_plot_do,
  ncol = 2,
  top = grid::textGrob(paste(climate, "-", optimal_value, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
)

@@

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]

# glimpse(biomass_list)
```

# Extract biomass values to match spatial and temporal range of climate vectors
```{r}
vocc_by_sp <- list()
for (i in seq_along(biomass_list)){

@@

if (length(variable_names) == 1) {
  vectors <- vocc %>%
    mutate(
      start_cell = icell, 
      vector_id = paste0(id, "_to_", tid),
      initial_value = var_1_s, 
      source_end_value = var_1_e, 
      target_value = as.numeric(target_values)
    ) %>%
    filter(distance > 1)  %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan)

  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))

} else {

  vectors <- vocc %>%
    mutate(
      start_cell = icell, vector_id = paste0(id, "_to_", tid),
      initial_value = s, source_end_value = e, target_value = target_values
    ) %>%
    filter(distance > 1) %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan)

  source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), function(x) data.frame(t(as.numeric(x))))
  vectors <- cbind(vectors, source_end)

  for (j in seq_along(variable_names)){
  vectors <- vectors %>% filter( paste0("X",j) < (optimal_value[j] - min_thresholds[j]) )
  }

@@

if (length(vectors[ ,2]) == 0) {
  try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})

  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$species <- spp 
  coefs$lifestage <- "NA"
  coefs$climate <- climate 
  coefs$min_thresholds <- min_thresholds
  coefs$max_thresholds <- max_thresholds
  coefs$optimum <- optimal_value
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  rbind(other_coefs, coefs)
  saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")

  stop("No vectors.")
}

@@

# combine source and target cell data
vocc_by_sp[[i]] <- rbind(vector_source, vector_target) %>% 
  filter(!is.na(icell)) %>% 
# remove potential source cells with no fish  
  #FIXME: should their be a flexible min density of fish for a source cell?
  filter(source_density > 0.000001)

# get mean and sd for densities without cell duplication
icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
mean.change <- mean(icells$perc_change, na.rm = TRUE)
sd.change <- sd(icells$perc_change, na.rm = TRUE)

@@

# calculate standardized density varibles
vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    perc_change = (end_density - start_density)/start_density, 
    cell_type = as.factor(cell_type),
    std_change = (perc_change)/sd.change, 
    std_source_density = (source_density - mean.density)/(sd.density*2), 
    mean_change = mean.change, sd_change = sd.change,
    mean_source_density = mean.density, sd_density = sd.density,
# set distance intercept at cell size
    distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, 
    distance2 = distance*distance) %>% 
  rename(X = x, Y = y)

if (length(variable_names) == 1) {
mean.climate <- mean(icells$end_climate, na.rm = TRUE)
sd.climate <- sd(icells$end_climate, na.rm = TRUE)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    mean_climate = mean.climate, sd_climate = sd.climate,
    std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
    std_end_climate2 = std_end_climate^2
    )

}else{

if (length(variable_names) == 2) {

vectors <- vectors %>% rename(
  source_end_climateA = X1, source_end_climateB = X2
  )

mean.climateA <- mean(icells$source_end_climateA, na.rm = TRUE)
sd.climateA <- sd(icells$source_end_climateA, na.rm = TRUE)
mean.climateB <- mean(icells$source_end_climateB, na.rm = TRUE)
sd.climateB <- sd(icells$source_end_climateB, na.rm = TRUE)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    mean_climateA = mean.climateA, sd_climate = sd.climateA,
    std_end_climateA = (source_end_climateA - mean.climateA)/(sd.climateA*2),
    std_end_climateA2 = std_end_climateA^2,
    mean_climateB = mean.climateB, sd_climate = sd.climateB,
    std_end_climateB = (source_end_climateB - mean.climateB)/(sd.climateB*2),
    std_end_climateB2 = std_end_climateB^2
    )
} else {
 stop("Doesn't currently support testing vectors bases on more than two climate variables at once.")
}
}

}
names(vocc_by_sp) <- names(biomass_list)

# glimpse(icells)
# range(icells$perc_change)
# range(vocc_by_sp[[i]]$std_change)
# range(vocc_by_sp[[i]]$X)

# glimpse(vocc_by_sp[[i]])
# trim spatial outliers 
# Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
# Xmin <- quantile(vocc_by_sp[[i]]$X, 0.005)
# std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
# 
# vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax )

# Adult

## Adult plots

@@

mature_plot2 <- ggplot(data, aes(x = distance, y = std_change, 
  group = cell_type, size = start_density, colour = end_climate)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y~x+I(x^2) )+
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_size_continuous(guide = "none") +
  # scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))

@@

mature_plot4 <- ggplot(filter(data, distance < 10), aes(source_density, y = std_change , 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", aes(group = cell_type)) +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

@@

## Adult models

### Regular mixed-model

@@

### Boosted regression
```{r}
data <- vocc_by_sp[["adult"]]

@@

### sdmTMB

```{r}
data <- vocc_by_sp[["adult"]]
# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$std_change, 0.9995)
# 
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

xy <- as.data.frame(cbind(data$X, data$Y))

  if (region == "Both odd year surveys") {
    n_knots <- 200
  }

  if (region == "West Coast Vancouver Island") {
    n_knots <- 100
  }

  # if (region == "West Coast Haida Gwaii") {
  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
  # }

if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(xy))}

  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = n_knots)
  sdmTMB::plot_spde(spde)

@@

# Immature

### Immature plots

@@

## Immature models

### Regular mixed-model
```{r}
idata <- vocc_by_sp[["imm"]]

# Xmax <- quantile(idata$X, 0.9995)
# Xmin <- quantile(idata$X, 0.005)
# std_max <- quantile(idata$std_change, 0.9995)
# 
# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )
#  
itest <- lmerTest::lmer(std_change ~ 
      cell_type * std_source_density + 
      cell_type * distance + cell_type * distance2 +
      std_end_climate + std_end_climate2 + (1 | vector_id), 
      REML = F, data = idata)
summary(itest)
plot(itest)
```
```{r eval=FALSE}
library(lme4)
itest3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
    cell_type * distance + cell_type * distance2 + 
    std_end_climate + std_end_climate2 + (1 | vector_id), data = idata)
summary(itest3)
plot(itest3, ask = FALSE)
```

### Boosted regression

@@

# Xmax <- quantile(idata$X, 0.9995)
# Xmin <- quantile(idata$X, 0.005)
# std_max <- quantile(idata$std_change, 0.9995)
# 
# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )

mi <- gbm::gbm(
  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
  data = idata, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )

@@

### sdmTMB

# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$std_change, 0.9995)
# 
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

ixy <- as.data.frame(cbind(idata$X, idata$Y))

  if (region == "Both odd year surveys") {
    n_knots <- 200
  }

  if (region == "West Coast Vancouver Island") {
    n_knots <- 100
  }

  # if (region == "West Coast Haida Gwaii") {
  #    n_knots <- 60
  # }

if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(ixy))}

  ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = n_knots)
  sdmTMB::plot_spde(ispde)

@@

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta","se")
coefs$ci <- coefs$se*1.96
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL

coefs <-cbind(var, coefs)
#coefs$var
#coefs$beta
#coefs$se
coefs$species <- spp 
coefs$lifestage <- "adult"
coefs$climate <- climate 
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_value
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

glimpse(coefs)
rm(other_coefs)
try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})

if (exists("other_coefs")) {
coefs <- rbind(other_coefs, coefs)
}
saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")

data$mean_source_density[1]
range(data$source_density)
```

```{r}
m <- itest

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta","se")
coefs$ci <- coefs$se*1.96
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL
coefs <-cbind(var, coefs)

coefs$species <- spp 
coefs$lifestage <- "immature"
coefs$climate <- climate 
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_value
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

glimpse(coefs)
try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
if (exists("other_coefs")) {
coefs <- rbind(other_coefs, coefs)
}
saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")

# intercept = mean std percent change in biomass for cell with avarage source biomass
idata$mean_source_density[1]
range(idata$source_density)
```
